---
title: "Curiosity and bare minimuming"
date: "2025-07-15T16:09"
categories: [FAQs]
toc-depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.align = "center",
  out.width = "90%",
  collapse = TRUE
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE
)
```

Hi everyone!

Remember that the core philosophy of my approach to teaching this class is to encourage you to be curious. Back in the FAQ post for session 2, [I said this about the ✓ grading system](/news/2025-06-18_faqs_session-02.qmd#why-do-you-use-s-for-grades):

---

There's a scene in season 1 of Ted Lasso where Ted cites a [made-up](https://www.snopes.com/fact-check/be-curious-not-judgmental-walt-whitman/) quote by Walt Whitman ([see timestamp 2:15 here](https://youtu.be/5x0PzUoJS-U?si=YSJvAoKo9UXx96I-&t=135)):

> Be curious, not judgmental.

Despite the fact that Walt Whitman never said this, I really like this sentiment, and I think it 100% applies to learning.

---

Especially now that we're out of the phase of the class with correct answers ([remember this!](/news/2025-07-09_no-more-answer-keys.qmd)), you have a lot of space to be curious. The assignment for session 9 involved making some sort of plot and adding annotations to it. In the assignment for session 11, you'll make some sort of dashboard.

Specifically, in exercise 9, the instructions said to add two each of these annotations to a plot of whatever data you wanted:

  - Text
  - An arrow
  - A rectangle

A few semesters ago, I got this for a submission for exercise 9:

```{r}
#| warning: false
#| message: false

library(tidyverse)

ggplot() +
  annotate(geom = "text", x = 1, y = 1, label = "Text") +
  annotate(geom = "text", x = 2, y = 1, label = "Text") +
  annotate(geom = "segment", x = 2.75, xend = 3.25, y = 1, yend = 1, arrow = arrow()) +
  annotate(geom = "segment", x = 3.75, xend = 4.25, y = 1, yend = 1, arrow = arrow()) +
  annotate(geom = "tile", x = 5, y = 1, width = 0.25, height = 0.25) +
  annotate(geom = "tile", x = 6, y = 1, width = 0.25, height = 0.25) 
```

With this as the description:

> Here are two labels, two arrows, and two rectangles. I completed the requirements. This assignment was useless.

And yeah, technically that meets the bare minimum requirements. And they learned a bit about `annotate()`. But I'm really not sure how much learning actually happened.

I received *a lot* of questions from you all last week via Discord and e-mail asking if your labels were right or said the same thing or if you had enough arrowheads or if three arrows could substitute as a label, etc.

Guess what—I don't actually care exactly how many annotations you added! At this point, we're in the curiosity phase of the class. **The whole point of every assignment in this course it to get you to play around with stuff. *That's* where the learning happens—learning happens through play and exploration.**

Please try to not bare minumum your exercises. I encourage you to take the time to *be curious* and play around with these assignments! Don't worry about exact answers!
